---
title: "Portfolio Assignment 2 – Introduction to Cultural Data Science"
author: "Emil Frej Brunbjeg"
date: "Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dslabs)
library(car)

```

## Part 1

*Load the 'divorce_margarine' dataset from the 'dslabs' package.
Investigate the correlation between margarine consumption and divorce
rates in Maine. Would an increase in the preference for margarine lead
to skyrocketing divorce rates?*

```{r }
# Load dataset
data("divorce_margarine", package = "dslabs")

#show
head(divorce_margarine)

```

The correlation between the variables is
`r cor(divorce_margarine$margarine_consumption)`. While this looks like
a perfect correlation, but correlation does not imply causation, and we
there cannot use this measure to say if an increase in the preference
for margarine lead to skyrocketing divorce rates.

## Part 2: Vocabulary Scores and Demographics

### Load and Filter the GSSvocab Dataset

```{r load-gssvocab}
# Load dataset


#filter year
GSSvocab <- GSSvocab %>%
  filter(year == 1978) %>%  # filter for year
  drop_na(everything()) # drop na in all col names



  



```

### Is Vocabulary Score Impacted by Education?

```{r vocab-education-model}
# Scatter plot of vocabulary score vs education
ggplot(gss_1978, aes(x = educ, y = vocab)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Vocabulary Score vs Education Level",
       x = "Education Level",
       y = "Vocabulary Score")

# Linear model
model_educ_vocab <- lm(vocab ~ educ, data = gss_1978)
summary(model_educ_vocab)
```

### Does Native-Born Status Affect Vocabulary Score?

```{r vocab-nativeborn-model}
# Boxplot for vocabulary score by native-born status
ggplot(gss_1978, aes(x = nativeBorn, y = vocab, fill = nativeBorn)) +
  geom_boxplot() +
  labs(title = "Vocabulary Score by Native-Born Status",
       x = "Native Born",
       y = "Vocabulary Score")

# Linear model with nativeBorn as predictor
model_native_vocab <- lm(vocab ~ educ + nativeBorn, data = gss_1978)
summary(model_native_vocab)
```

### Relationship Between Education Level and Native-Born Status

```{r education-nativeborn-relationship}
# Visualize relationship between education and native-born status
ggplot(gss_1978, aes(x = nativeBorn, y = educ, fill = nativeBorn)) +
  geom_boxplot() +
  labs(title = "Education Level by Native-Born Status",
       x = "Native Born",
       y = "Education Level")

# Linear model with interaction term
model_interaction <- lm(vocab ~ educ * nativeBorn, data = gss_1978)
summary(model_interaction)
```

### Model Comparison

```{r model-comparison}
# Compare model summaries
summary(model_educ_vocab)
summary(model_native_vocab)
summary(model_interaction)
```

This R Markdown file analyzes the relationship between margarine
consumption and divorce rates in Maine, then explores the impact of
education and native-born status on vocabulary scores.
